Characterizing edges in signed and vector-valued graphs

### 2017-12-17 troll/algs_batch.tex
pass over troll

Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
labeled edges from $E(Y)$. Formally, we have a training set $\trainset
= \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, that
has been drawn from $E\times Y$ \uar{} with replacement, with
$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
 We want to build a predictive model for the labels of the remaining
edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
locally the parameters by their empirical means in the training set, which under some density
assumptions are showed to concentrate around their true values. The second algorithm, on the other
hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
parameters through a global label propagation approach, without making any density assumption.

@@

We are now ready to describe our algorithm, which follows naturally from the formula derived above.
It takes as input a training set \trainset{} drawn at random with replacement and outputs a sign
prediction for all the edges in $E\setminus \trainset$. We call it \usrule{}, which stands for Bayes
Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness.
\iffalse
Despite its apparent simplicity, \usrule{}
works reasonably well in practice, as demonstrated by our experiments (see \autoref{s:exp}).
Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes
with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier
$y^*(u,v)$ with respect to the generative model defined in \autoref{s:gen}. It is important to point
out that when we use $1-\htr(u)$ and $1-\hun(v)$ to estimate $p_u$ and $q_v$, an additive bias shows
up due to \eqref{e:pout} and \eqref{e:pin}. This motivates the need of a threshold parameter $\tau$
to cancel this bias. Yet, the presence of a prior distribution $\mu(p,q)$ ensures that this bias is
the same for all edges $(u,v) \in E$.

@@

Remember we suppose that the training set \trainset{} has been drawn \uar{} with replacement, with
$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
likelihood estimator of the parameters $\{p_u, q_u\}_{u=1}^{|V|}$ based on \trainset{}.
If we further assume, in order to make the computation more tractable, that
the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
independent marginals,\footnote{As we will see, in real data, around $80\%$ if the edges are
positive, meaning this assumption of uniformity over $[0,1]^2$ is unlikely to fully holds, for
otherwise the signs would more balanced}
we show in the supplementary material \vpageref{ssec:troll_mle_deriv} that for $\ell \in \{1,
\ldots, |V|\}$ the gradient of the log-likelihood function with respect to a given $p_\ell$ and
$q_\ell$ satisfies

@@

The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} +
f_{E\setminus E_0}$ in \eqref{e:quadratic}, where the regularization term tries to enforce the extra
constraint that whenever a node $u$ has a high out-degree then the corresponding $p_u$ should be
close to \shalf. Thus, on any edge $(u,v)$ departing from $u$, the Bayes optimal predictor $y^*(u,v)
= \sgn(p_u+q_v-1)$ will mainly depend on $q_v$ being larger or smaller than \shalf{} (assuming $v$
has small in-degree). Similarly, if $u$ has a high in-degree, then the corresponding $q_u$ should be
close to \shalf{} implying that on any edge $(v,u)$ arriving at $u$ the Bayes optimal predictor
$y^*(v,u)$ will mainly depend on $p_v$ (assuming $v$ has small out-degree). Put differently, a node
having a huge out-neighborhood makes each outgoing edge \enquote{count less} than a node having only
a small number of outgoing edges, and similarly for in-neighborhoods.

>>>>

Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
labeled edges from $E(Y)$. Formally, we have a training set $\trainset
= \big((u_1,v_1),y_{u_1,v_1}), ((u_2,v_2),y_{u_2,v_2}), \ldots, ((u_m,v_m),y_{u_m,v_m}\big)$, that
has been drawn from $E\times Y$ \uar{} without replacement, with
$m = |E_0|$. We want to build a predictive model for the labels of the remaining
edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
locally the parameters by their empirical means in the training set, which under some density
assumptions are showed to concentrate around their true values. The second algorithm, on the other
hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
parameters through a global label propagation approach, without making any density assumption.

@@

We are now ready to describe our algorithm, which follows naturally from the formula derived above.
It takes as input a training set \trainset{} drawn at random without replacement and outputs a sign
prediction for all the edges in $E\setminus \trainset$. We call it \usrule{}, which stands for Bayes
Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness.
\iffalse
Despite its apparent simplicity, \usrule{}
works reasonably well in practice, as demonstrated by our experiments (see \autoref{s:exp}).
Moreover, unlike previous edge sign prediction methods for directed graphs, our classifier comes
with a rigorous theoretical motivation, since it approximates the Bayes optimal classifier
$y^*(u,v)$ with respect to the generative model defined in \autoref{s:gen}. It is important to point
out that when we use $1-\htr(u)$ and $1-\hun(v)$ to estimate $p_u$ and $q_v$, an additive bias shows
up due to \eqref{e:pout} and \eqref{e:pin}. This motivates the need of a threshold parameter $\tau$
to cancel this bias. Yet, the presence of a prior distribution $\mu(p,q)$ ensures that this bias is
the same for all edges $(u,v) \in E$.

@@

Remember we suppose that the training set \trainset{} has been drawn \uar{} without replacement, with
$m = |E_0|$. Then a reasonable approach to approximate $y^*(u,v)$ would be to resort to a maximum
likelihood estimator of the parameters $\{p_u, q_u\}_{u=1}^{|V|}$ based on \trainset{}.
If we further assume, in order to make the computation more tractable, that
the joint prior distribution $\mu(p,q)$ is uniform over $[0,1]^2$ with
independent marginals,\footnote{As we will see, in real data, around $80\%$ if the edges are
positive, meaning this assumption of uniformity over $[0,1]^2$ is unlikely to fully holds, for
otherwise the signs would more balanced.}
we show in the supplementary material \vpageref{ssec:troll_mle_deriv} that for $\ell \in \{1,
\ldots, |V|\}$ the gradient of the log-likelihood function with respect to a given $p_\ell$ and
$q_\ell$ satisfies

@@

The function ${\widehat f}$ is thus a regularized version of the target function $f_{E_0} +
f_{E\setminus E_0}$ in \eqref{e:quadratic}, where the regularization term tries to enforce the extra
constraint that whenever a node $u$ has a high out-degree then the corresponding $p_u$ should be
close to \shalf. Thus, on any edge $(u,v)$ departing from $u$, the Bayes optimal predictor $y^*(u,v)
= \sgn(p_u+q_v-1)$ will mainly depend on $q_v$ being larger or smaller than \shalf{} (assuming $v$
has small in-degree). Similarly, if $u$ has a high in-degree, then the corresponding $q_u$ should be
close to \shalf{}, implying that on any edge $(v,u)$ arriving at $u$, the Bayes optimal predictor
$y^*(v,u)$ will mainly depend on $p_v$ (assuming $v$ has small out-degree). Put differently, a node
having a huge out-neighborhood makes each outgoing edge \enquote{count less} than a node having only
a small number of outgoing edges, and similarly for in-neighborhoods.

### 2017-12-17 troll/experiments.tex
finish troll layout
pass over troll

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
technique using a list of positive and negative words; see
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

\begin{table}[bt]
  \centering
  \small
  \caption{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
  columns provide two different measures of label regularity, while the last two columns give the
  proportion of reciprocal edges, and among them the fraction with different signs.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
    \thead{reciprocal\\ disagreement} \\
    \midrule                                                                                                   
    \aut{}  & \np{4831}   & \np{39452}  & 8.2               & 72.3\%              & .076                          & .191                    & 5.1\%      & 27.1\%            \\
    \adv{}  & \np{5417}   & \np{51312}  & 9.5               & 75.1\%              & .061                          & .132                    & 33.6\%     & 28.6\%            \\
    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.8\%              & .063                          & .142                    & 5.6\%      & 10.0\%            \\
    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.4\%              & .059                          & .143                    & 17.7\%     & 4.0\%             \\
    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.3\%              & .031                          & .074                    & 30.8\%     & 2.1\%             \\
    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.9\%              & .034                          & .086                    & 6.5\%      & 14.6\%            \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729

@@

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the initial $p'$ value of the corresponding node. Then, because we want to model real
sign distribution and have more
positive than negative edges, we apply an exponential transformation \[ p = \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p'} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}\footnote{More precisely,
these figures are normalized histogram, meaning the counts are scale down so that the bars total
area sums up to one.}, giving $71.9\%$ of positive edges.

% TODO: do a proper subfigure instead of minipage?
\vspace{\baselineskip}
\noindent\begin{minipage}{187mm} 
  \begin{minipage}{46mm}
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
    \captionof{subfigure}{Density of $p$.}
    % \label{fig:figure1}
  \end{minipage}%
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf} 
    \captionof{subfigure}{Density of $q$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf} 
    \captionof{subfigure}{Density of $\eta$.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \begin{minipage}{46mm}
    \centering 
    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
    \captionof{subfigure}{Density of $\eta$ for both signs.} 
    % \label{fig:figure2} 
  \end{minipage} 
  \captionof{figure}{Synthetic distributions on \wik{}.} 
  \label{fig:troll_synth_distrib} 
\end{minipage}
\vspace{\baselineskip}

@@

The result of this comparison on the six datasets is showed in \autoref{tab:troll_synth_global}.
For each network, we generated the signs once. Then, for different training size (20\%, 40\% or 80\%
of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
predictor) or its estimated value (\usrule{}).
The MCC of the Bayes predictor is the same on all datasets (around $30$\footnote{Here and in the
following, we multiply all MCC value by 100 to improve readability.}), and so is its accuracy
(around $75\%$). The corresponding values for \usrule{} are close on all datasets (although some are
easier) and the gap naturally decreases as the training size increases. Another interesting
quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and then specifically for
the testing edges whose both endpoints have been sampled above a certain threshold in the training
set. As expected, the estimation is more accurate with increasing training set size, and with
increasing number of samples for a given edge.

\input{synth_res}

@@

  \item The \usrule{} algorithm described in \autoref{ss:passive}.

@@

  \item  The solution obtained by directly solving the unregularized problem \eqref{e:quadratic}
    through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
    binarizing threshold was set by cross-validation on the training set.\footnote{We have also
    tried to minimize \eqref{e:quadratic} by removing the $[-1,+1]$ constraints, but got similar MCC
    results as the ones we report for \qoptim{}}

@@

Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
training on sets of varying size (from 5\% to 25\%, plus 50\% and 90\%). Results have been averaged
over 12 repetitions.
Because scalability is a major concern when training on sizeable datasets, we also give an idea of
relative training times (in milliseconds) by reporting separately in \autoref{tab:all_time} the time it took to train a single run of
each algorithm on a training set of size\footnote{Comparison of training time performances is fair
since all algorithms have been carefully implemented using the same stack of Python libraries, and
run on the same machine (16 Xeon cores and 192Gb Ram).} 15\% of $|E|$, and then predict on the test
set. Some conclusions emerge from those experiments:

@@

The explanation of
why these two sampling strategies produce so different results can be inferred from
\autoref{tab:troll_early_sampling}. Letting $V_{\mathrm{out}}$ be the subset of node of $V$ with at
least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
naturally larger when sampling at random. Yet, when the training set reaches 90\% size, both
sampling strategies cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
$V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for
testing edges \euv{}, the breakdown between three situations, from the least to the most
informative:
\begin{enumerate}[(i),nosep]
  \item no sampled edge outgoing from $u$ nor incoming to $v$;
  \item some sampled edges, but only either outgoing from $u$ or incoming to $v$; and
  \item some sampled edges, both outgoing from $u$ and incoming to $v$;
\end{enumerate}
When the training set is build \uar{}, the vast majority of testing edges falls into the case (iii).
This is no longer true with temporal training set, where situations (ii) and even (i) are more
common, making prediction more difficult. The difference of repartition is particularly strong in
\wik{}, justifying the better performance in \epi{}.

>>>>

We consider six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}.
In \adv{}, a trust-based social network for open source developers, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric, and its
resistance to attacks, is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
technique using a list of positive and negative words; see
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

\begin{table}[bt]
  \centering
  \small
  \caption[\dssn{} dataset properties]{Dataset properties. The 5\thup{} column gives the fraction of positive labels. The next two
  columns provide two different measures of label regularity, while the last two columns give the
  proportion of reciprocal edges, and among them the fraction with different signs.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
    \thead{reciprocal\\ disagreement} \\
    \midrule                                                                                                   
    \aut{}  & \np{4831}   & \np{39452}  & 8.2               & 72.3\%              & .076                          & .191                    & 5.1\%      & 27.1\%            \\
    \adv{}  & \np{5417}   & \np{51312}  & 9.5               & 75.1\%              & .061                          & .132                    & 33.6\%     & 28.6\%            \\
    \wik{}  & \np{7114}   & \np{103108} & 14.5              & 78.8\%              & .063                          & .142                    & 5.6\%      & 10.0\%            \\
    \sla{}  & \np{82140}  & \np{549202} & 6.7               & 77.4\%              & .059                          & .143                    & 17.7\%     & 4.0\%             \\
    \epi{}  & \np{131580} & \np{840799} & 6.4               & 85.3\%              & .031                          & .074                    & 30.8\%     & 2.1\%             \\
    \kiw{}  & \np{138587} & \np{740106} & 5.3               & 87.9\%              & .034                          & .086                    & 6.5\%      & 14.6\%            \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Finally, there is a low fraction of reciprocal edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases, such reciprocal edges do not disagree,
\ie{} they have the same sign. In practice, we can use this fact to improve our prediction at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\yhat_{u,v} = y_{v,u}$. For clarity, when comparing
our methods with existing approaches, we dot not use that heuristic. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729

@@

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has a parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior, such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edge, and likewise the in-degree of nodes with at least one incoming edge. For
that, in the case of $p$, we partition the interval $[0,1]$ into $|V|$ segments of size proportional
to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the initial $p'$ value of the corresponding node. Then, because we want to model real
sign distribution and have more
positive than negative edges, we apply an exponential transformation \[ p = \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p'} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}\footnote{More precisely,
these figures are normalized histograms, meaning the counts are scaled down so that the bars total
area sums up to one.}, giving $71.9\%$ of positive edges.

\begin{figure*}[hbtp]
  \centering
  \begin{subfigure}[b]{0.401\textwidth} 
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_pdense.pdf}
    \caption{Density of $p$.}
  \end{subfigure}\quad
  \begin{subfigure}[b]{0.401\textwidth} 
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_qdense.pdf}
    \caption{Density of $q$.}
  \end{subfigure}

  \begin{subfigure}[b]{0.401\textwidth} 
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_etadense.pdf}
    \caption{Density of $\eta$.}
  \end{subfigure}\quad
  \begin{subfigure}[b]{0.401\textwidth} 
    \centering
    \includegraphics[width=.95\linewidth]{raw/synth_wik_signdense.pdf}
    \caption{Density of $\eta$ for both signs.}
  \end{subfigure}
  \caption{Synthetic distributions on \wik{}.} 
  \label{fig:troll_synth_distrib} 
\end{figure*}

@@

\input{synth_res}
The result of this comparison on the six datasets is showed in \autoref{tab:troll_synth_global}.
For each network, we generated the signs once. Then, for different training size (20\%, 40\% or 80\%
of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
predictor) or its estimated value (\usrule{}).
The MCC of the Bayes predictor is the same on all datasets (around $30$\footnote{Here and in the
following, we multiply all MCC value by 100 to improve readability.}), and so is its accuracy
(around $75\%$). The corresponding values for \usrule{} are close on all datasets (although some are
easier) and the gap naturally decreases as the training size increases. Another interesting
quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and then specifically for
the testing edges whose both endpoints have been sampled above a certain threshold in the training
set. As expected, the estimation is more accurate with increasing training set size, and with
increasing number of samples for a given edge.

@@

  \item The \usrule{} algorithm described in \autoref{ss:bayes_approx}.

@@

  \item  The solution obtained by directly solving the unregularized problem \eqref{e:quadratic}
    through a constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
    binarizing threshold was set by cross-validation on the training set.\footnote{We have also
    tried to minimize \eqref{e:quadratic} by removing the $[-1,+1]$ constraints, but got similar MCC
    results as the ones we report for \qoptim{}}

@@

Our main results are summarized in \autoref{tab:all_mcc}, reporting MCC test set performance after
training on sets of varying size (from 5\% to 25\%, plus 50\% and 90\%). Results have been averaged
over 12 repetitions.
Because scalability is a major concern when training on sizeable datasets, we also give an idea of
relative training times by reporting separately in \autoref{tab:all_time} the time (in milliseconds) it took to train a single run of
each algorithm on a training set of size\footnote{Comparison of training time performances is fair
since all algorithms have been carefully implemented using the same stack of Python libraries, and
run on the same machine (16 Xeon cores and 192GB Ram).} 15\% of $|E|$, and then predict on the test
set. Some conclusions emerge from those experiments:

@@

The explanation of
why these two sampling strategies produce so different results can be inferred from
\autoref{tab:troll_early_sampling}. Letting $V_{\mathrm{out}}$ be the subset of node of $V$ with at
least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
naturally larger when sampling at random. Yet, when the training set reaches 90\% size, both
sampling strategies cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
$V_{\mathrm{in}}$. The difference in MCC is thus explained by the last three columns. It shows, for
testing edges \euv{}, the breakdown between three situations, from the least to the most
informative:
\begin{enumerate}[(i),nosep]
  \item no sampled edge outgoing from $u$ nor incoming to $v$;
  \item some sampled edges, but only either outgoing from $u$ or incoming to $v$; and
  \item some sampled edges, both outgoing from $u$ and incoming to $v$;
\end{enumerate}
When the training set is build \uar{}, the vast majority of testing edges falls into the case (iii).
This is no longer true with temporal training set, where situations (ii) and even (i) are more
common, making prediction more difficult. The difference of repartition is particularly strong in
\wik{}, justifying the better performance in \epi{}.
